---
title: "Airbnb in NYC with SparkR and Sparkling water"
author: "Dominik Klepl"
date: "2/6/2020"
output: html_document
---

## Connect to Spark
Load required libraries
```{r echo=FALSE}
library(sparklyr)
library(rsparkling)
library(dplyr)
library(ggplot2) #for plots
library(ggthemes)
library(magrittr) #pipes support
```

Launch Spark session and create Spark context (sc)
```{r}
sc = spark_connect(master = "local", version = "2.2.1")
```

Create H2O context
```{r}
#hc = H2OContext.getOrCreate(sc)
```

Open H2O (Sparkling water) UI
```{r}
#hc$openFlow()
```

### Load data
Our dataset is a csv file. Now we'll load into spark.
```{r}
#path to dataset
DATA_PATH = "data/AB_NYC_2019.csv"
data = spark_read_csv(sc = sc, name = "data", path = DATA_PATH)
class(data)
```

Let's look at the data
```{r}
cat("The dataset has", sdf_nrow(data), "datapoints","\nand", sdf_ncol(data), "features.")
```

What columns do we have?
```{r}
colnames(data)
```

Some of these columns won't be very useful in the further analysis so we exclude them.
```{r}
data = data %>%
  select(-c(id, host_name))
```


## Missing values
First, we'll deal with the missing values and try to avoid dropping them as much as possible.
```{r}
(missing_values = data %>% 
  mutate_all(is.na) %>%
  mutate_all(as.numeric) %>%
  summarise_all(sum) %>% 
  collect() %>%
  t() %>%
  as.data.frame())
  
colnames(missing_values) = "NAs"
missing_values$Feature = rownames(missing_values)
rownames(missing_values) = NULL

missing_values = missing_values %>% filter(NAs !=0)

(NA_plot = ggplot(missing_values, aes(x = Feature, y = NAs, fill=Feature))+
  geom_col()+
  theme_few()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())+
  labs(title = "",
       x = "",
       y = "Sum of missing values"))
```

Most missing values are in last_review and reviews_per_month. This probably happens when a listing has no reviews yet. Let's test that hypothesis.
```{r}
#how many number_of_reviews==0 have NA in reviews_per_month 
NA_test = data %>% 
  select(number_of_reviews, reviews_per_month, last_review) %>%
  filter(number_of_reviews==0) %>%
  mutate(na_month = as.character(is.na(reviews_per_month))) %>%
  mutate(na_last = as.character(is.na(last_review)))

NA_test %>%
  group_by(na_month) %>%
  summarise(count = n())

NA_test %>%
  group_by(na_last) %>%
  summarise(count = n())
```

Most of NAs in last_review and reviews_per_month were indeed the result of having no reviews yet. We can therefore replace these NAs with 0.
There are also a few missing values in name and host_id, we'll replace these with "unknown" as there might be a use for that column.
Since host_id and host_name contain the same information, we'll exclude the host_name.
```{r}
data = data %>%
  #replace with 0
  mutate(last_review = ifelse(is.na(last_review), 0, last_review)) %>% 
  #replace with 0
  mutate(reviews_per_month = ifelse(is.na(reviews_per_month), 0, reviews_per_month)) %>% 
  #replace with unknown
  mutate(name = ifelse(is.na(name), "Unknown", name)) %>%
  #replace with unknown
  mutate(host_id = ifelse(is.na(host_id), "Unknown", host_id))
```



