---
title: "02 Explorative analysis and visualisations"
author: "Dominik Klepl"
date: "2/7/2020"
output: html_document
---

## Load libraries, connect Spark, load data
```{r setup, include=FALSE}
library(sparklyr)
library(rsparkling)
library(dplyr)
library(ggplot2) #for plots
library(ggthemes)
library(magrittr) #pipes support
library(gridExtra) #creating grids of plots
library(patchwork) #more complicated plot layouts
library(viridis)
library(ggmap)

sc = spark_connect(master = "local", version = "2.2.1")

VISUALIZATIONS = "documentation/figures"
d1 = spark_read_csv(sc = sc, path = "data/cleaned_data.csv")
d1 = d1 %>% filter(neighbourhood !="1464358")
```

Turn columns into the right format
```{r}
d1 = d1 %>% 
  mutate(host_id = as.numeric(host_id)) %>%
  mutate(latitude = as.numeric(latitude)) %>%
  mutate(longitude = as.numeric(longitude)) %>%
  mutate(price = as.numeric(price)) %>%
  mutate(minimum_nights = as.numeric(minimum_nights)) %>%
  mutate(number_of_reviews = as.numeric(number_of_reviews)) %>%
  mutate(reviews_per_month = as.numeric(reviews_per_month)) %>%
  mutate(calculated_host_listings_count = as.numeric(calculated_host_listings_count)) %>%
  mutate(availability_365 = as.numeric(availability_365)) %>%
  na.omit() %>%
  rename(desc = name,ID = host_id, listing_host = calculated_host_listings_count, availability = availability_365)

length = sdf_nrow(d1)
```


# Single variable exploration - Hunting outliers

## Distribution
We'll start by plotting distribution of all numerical variables.
```{r}
#histogram
plot_histogram = function(col, data = d1) {
  plot_data = data %>% sdf_read_column(col) %>% as.numeric() %>% as.data.frame()
  colnames(plot_data) = col
  
  plot = ggplot(plot_data, aes_string(x = col))+
  geom_histogram(bins = 30, fill="#1F77B4")+
  theme_few()
  labs(title = col,
       x = "",
       y = "")+
  theme(plot.title = element_text(hjust = 0.5, size=9),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
  
  return(plot)
}

plot_box = function(col, data = d1) {
  plot_data = data %>% sdf_read_column(col) %>% as.numeric() %>% as.data.frame()
  colnames(plot_data) = col
  
  plot = ggplot(plot_data, aes_string(y = col))+
    geom_boxplot()+
    theme_few()+
    labs(title = col,
         x = "",
         y = "")+
    theme(plot.title = element_text(hjust = 0.5, size=9))
  
  return(plot)
}
```

Loop over numeric columns and plot histograms
```{r}
numeric_columns = colnames(d1)[c(5,6,8,9,10,12:14)]

histograms = list()
for (i in 1:length(numeric_columns)){
  hist = plot_histogram(numeric_columns[i])
  histograms[[numeric_columns[i]]] = hist
}

#arrange histograms in 2x4 grid
do.call("grid.arrange", c(histograms, ncol = 2))

#save the grid as png
hist_grid = do.call("arrangeGrob", c(histograms, ncol = 2))
ggsave(paste0(VISUALIZATIONS, "/histogram_grid.png"), hist_grid, width = 7, height = 7)
```

## Boxplots
```{r}
boxplots = list()
for (i in 1:length(numeric_columns)){
  box = plot_box(numeric_columns[i])
  boxplots[[numeric_columns[i]]] = box
}

#arrange histograms in 2x4 grid
do.call("grid.arrange", c(boxplots, ncol = 2))

#save the grid as png
box_grid = do.call("arrangeGrob", c(boxplots, ncol = 2))
ggsave(paste0(VISUALIZATIONS, "/boxplot_grid.png"), box_grid)
```

### Availability
We'll start by looking for nonsense values (e.g. availability more than 365) and explore whether it needs transformations, excluding outliers etc.
```{r}
d1 %>%
  sdf_describe("availability")

#look at its histogram in detail
histograms$availability
boxplots$availability

#there is at least one listing that is never available
cat("How many accomodations are never available?",
    d1 %>% filter(availability==0) %>% sdf_nrow())

#on the other hand, how many are always available
cat("\nHow many accomodations are always available?",
    d1 %>% filter(availability==365) %>% sdf_nrow())
```

It seems most listing are available only for very narrow time windows. However, there is quite high variation and so the mean and standard deviations are shifted away from the mode.
Altough log-transform might help to achieve normal distribution, we'd loose data as there are many 0 values (log(0)=infinity).


### Minimum nights
```{r}
d1 %>%
  sdf_describe("minimum_nights")

#look at histogram and boxplot
histograms$minimum_nights
boxplots$minimum_nights

cat("How many accomodations have minimum nights a year or more?",
    d1 %>% filter(minimum_nights >= 365) %>% sdf_nrow())

cat("How many accomodations have minimum nights more than mean + 2.5*standard deviation?",
d1 %>% filter(minimum_nights >= mean(minimum_nights, na.rm = T)+2.5*sd(minimum_nights, na.rm = T)) %>% sdf_nrow())
```

There are 427 datapoints that are more than 2.5 SDs away from mean, it's safe to say that these are outliers, we'll remove them later for modelling purposes.
```{r}
#d1 = d1 %>%
#  filter(minimum_nights <= mean(minimum_nights, na.rm = T)+2.5*sd(minimum_nights, na.rm = T))

#replot histogram
#plot_histogram("minimum_nights")
#plot_box("minimum_nights")
```
There are still some possible outliers but the spread is significantly lower. Regression models should be able to handle extrapolating these more extreme values.

## Number of reviews
```{r}
d1 %>% sdf_describe("number_of_reviews")

#look at plots
histograms$number_of_reviews
boxplots$number_of_reviews
```

We can see some extreme values again.
```{r}
d1 %>%
  filter(number_of_reviews >= mean(number_of_reviews, na.rm = T)+2.5*sd(number_of_reviews, na.rm = T)) %>%
  sdf_nrow()

d1 %>%
  filter(number_of_reviews == 0) %>%
  sdf_nrow()
```


## Reviews per month
```{r}
d1 %>% sdf_describe("reviews_per_month")

histograms$reviews_per_month
boxplots$reviews_per_month
```

Most listings receive approx 1 review/month with a few extremes.

## Listings per host + host_ID
```{r}
d1 %>% sdf_describe("listing_host")

histograms$listing_host
```

This feature is a bit misleading since there are duplicates (listings from the same host). Let's recompute the statistics by looking at unique values.
```{r}
listings = d1 %>% select(ID, listing_host) %>% distinct() %>% arrange(desc(listing_host))

listings %>% sdf_describe("listing_host")

#replot histogram and boxplot
(hist_listings = plot_histogram("listing_host", listings))
(box_listings = plot_box("listing_host", listings))

#update grids of plots
histograms[["listing_host"]] = hist_listings
boxplots[["listing_host"]] = box_listings

#save to png again
box_grid = do.call("arrangeGrob", c(boxplots, ncol = 2))
ggsave(paste0(VISUALIZATIONS, "/boxplot_grid.png"), box_grid)
hist_grid = do.call("arrangeGrob", c(histograms, ncol = 2))
ggsave(paste0(VISUALIZATIONS, "/histogram_grid.png"), hist_grid, width = 7, height = 7)
```

Most hosts have only 1 listing on average. However, there are clearly some who have sort of a network of listings. Let's look at the top 10.
```{r}
#collect listings to R
listing_data  = sdf_collect(listings)
listing_data$ID = as.factor(listing_data$ID)

#plot trend of owning airbnb
(top_hosts = ggplot(listing_data[1:10,], aes(x = reorder(ID, -listing_host), y = listing_host, fill = ID))+
  geom_bar(stat = "identity")+
  theme_few()+
  scale_fill_tableau()+
  guides(fill = F)+
  labs(x = "",
       y = "Listings")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)))
ggsave(paste0(VISUALIZATIONS, "/top_hosts_listings.png"), top_hosts)

#IDs for later with more than 10
topID = listing_data %>% filter(listing_host>=10) %>% select(ID)
topID = as.numeric(as.character(topID$ID))
``` 

## Price
```{r}
d1 %>% sdf_describe("price")

histograms$price
boxplots$price
```

There are some ridiculously expensive listings. On the other hand, there's at least one listing which is offered for free.
To explore the distribution of more common prices, we'll remove these extremes for now.
```{r}
cat("There are", d1 %>% filter(price == 0) %>% sdf_nrow(), "listings offered for free.")

#Let's at their descriptions
d1 %>% filter(price == 0) %>% select(desc) #these mostly look like it's a some invalid data

#remove those free places immediately
d1 = d1 %>% filter(price > 0)

#Let's look at descriptions of the most expensive listings
d1 %>% filter(price > 5000) %>% select(price,desc) %>% arrange(desc(price))

prices = d1 %>% filter(price < 600) 

plot_histogram("price", prices)
```

## Categorical variables

Summarise the categorical columns - count values per category
```{r}
(neigbourhood_gr_count = d1 %>%
   group_by(neighbourhood_group) %>%
   summarise(count = n()) %>%
   arrange(desc(count)) %>%
   rename(borough = neighbourhood_group) %>%
   sdf_collect() %>%
   as.data.frame())

neigbourhood_count = d1 %>%
  group_by(neighbourhood) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  sdf_collect() %>%
  as.data.frame()
neigbourhood_count[1:10,]

(room_count = d1 %>%
  group_by(room_type) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  sdf_collect() %>%
  as.data.frame())
```

Plot the counts
```{r}
categ1 = ggplot(neigbourhood_gr_count, aes(x = reorder(borough, -count), y = count, fill = borough))+
  geom_bar(stat = "identity")+
  theme_few()+
  scale_fill_tableau()+
  labs(x = "")+
  guides(fill = F)

categ2 = ggplot(neigbourhood_count[1:10,], aes(x = reorder(neighbourhood, -count), y = count, fill = neighbourhood))+
  geom_bar(stat = "identity")+
  theme_few()+
  scale_fill_tableau()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(x = "")+
  guides(fill = F)

categ3 = ggplot(room_count, aes(x = reorder(room_type, -count), y = count, fill = room_type))+
  geom_bar(stat = "identity")+
  theme_few()+
  scale_fill_tableau()+
  labs(x = "")+
  guides(fill = F)

#arrange plots in nice layout
(categorical_counts = (categ1 + categ3)/categ2)

ggsave(paste0(VISUALIZATIONS, "/categorical_counts.png"), categorical_counts)
```

# Price vs. rest
Becaue we're mainly interested in prices, we'll now explore the relationship of price and other variables.

### Price vs. Borough
```{r}
price_bor = d1 %>%
   group_by(neighbourhood_group) %>%
   summarise(mean = mean(price),
             sd = sd(price)) %>%
  sdf_collect()

(price_bor_bar = ggplot(price_bor, aes(x = reorder(neighbourhood_group, -mean), y = mean, fill = neighbourhood_group))+
  geom_col(color = "black", size = 0.3)+
  geom_errorbar(aes(ymin =mean-0.5*sd, ymax = mean+0.5*sd))+
  theme_few()+
  scale_fill_tableau()+
  guides(fill = F)+
  labs(x = "",
       y = "price")+
  theme(axis.text.x = element_text(angle = 30, hjust = 1)))

ggsave(paste0(VISUALIZATIONS, "/price_borough.png"), price_bor_bar)
```

### Price vs. Neighbourhood
```{r}
price_nei = d1 %>%
  group_by(neighbourhood) %>%
  summarise(mean = mean(price),
            sd = sd(price)) %>%
  arrange(desc(mean)) %>%
  sdf_collect()

(price_nei_point = ggplot(price_nei, aes(x = reorder(neighbourhood, -mean), y = mean))+
  geom_point()+
  theme_few()+
  labs(x = "",
       y = "price"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 4))

(price_nei_bar = ggplot(price_nei[1:10,], aes(x = reorder(neighbourhood, -mean), y = mean, fill = neighbourhood))+
  geom_col(color = "black", size = 0.3)+
  geom_errorbar(aes(ymin =mean-0.5*sd, ymax = mean+0.5*sd), width = 0.5)+
  theme_few()+
  scale_fill_tableau()+
  guides(fill = F)+
  labs(x = "",
       y = "price")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)))

ggsave(paste0(VISUALIZATIONS, "/price_neighbourhoood_all.png"), price_nei_point)
ggsave(paste0(VISUALIZATIONS, "/price_neighbourhoood_top10.png"), price_nei_bar)
```

### Price vs. Room type
```{r}
(price_type = d1 %>%
  group_by(room_type) %>%
  summarise(mean = mean(price),
            sd = sd(price)) %>%
  arrange(desc(mean)) %>%
  sdf_collect())

(price_type_bar = ggplot(price_type, aes(x = reorder(room_type, -mean), y = mean, fill = room_type))+
  geom_col(color = "black", size = 0.3)+
  geom_errorbar(aes(ymin =mean-0.5*sd, ymax = mean+0.5*sd), width = 0.5)+
  theme_few()+
  scale_fill_tableau()+
  guides(fill = F)+
  labs(x = "",
       y = "price")+
  theme(axis.text.x = element_text(angle = 30, hjust = 1)))
```

## Price vs Borough/Room type
Let's look whether there is interaction between borough and room type that explains the price variations.
```{r}
borough_type = d1 %>%
  mutate(borough_type = paste(neighbourhood_group, room_type, sep = "-")) %>%
  group_by(borough_type) %>%
  summarise(mean = mean(price),
            sd = sd(price)) %>%
  sdf_collect() %>%
  tidyr::separate(borough_type, c("borough", "room_type"), sep = "-")

(borough_type_bar = ggplot(borough_type, aes(fill=room_type, y=mean, x=borough))+
  geom_col(position="dodge", color = "black", size = 0.3)+
  geom_errorbar(aes(ymin =mean-0.5*sd, ymax = mean+0.5*sd), position = position_dodge(0.9), width = 0.5)+
  theme_few()+
  scale_fill_tableau(name = "Room type")+
  labs(x = "",
       y = "price"))
```

#### Borough/type vs price layout
```{r}
(bor_type_layout = (price_bor_bar + price_type_bar) / borough_type_bar)
ggsave(paste0(VISUALIZATIONS, "/borough_type.png"), bor_type_layout, width = 7, height = 6)
```

## Maps
First, we create a background map of New York City + some detailed maps of boroughs
```{r}
register_google(key = "XXX")


#New York
lat = d1 %>% summarise(lat = mean(latitude)) %>% sdf_collect %>% as.numeric()
lon = d1 %>% summarise(lon = mean(longitude)) %>% sdf_collect %>% as.numeric()

nyc = get_googlemap(center = c(lon = lon, lat = lat),
                     zoom = 11, scale = 2,
                     maptype = "roadmap",
                     color = "color")
```

### Price map
```{r}
price_data = d1 %>% 
  select(latitude, longitude, price, neighbourhood_group) %>%
  filter(price <500)

(price_map = ggmap(nyc)+
  geom_point(data = price_data,aes(x = longitude, y = latitude, color = price), alpha = 0.8)+
  theme_map()+
  scale_color_gradient2_tableau())

ggsave(paste0(VISUALIZATIONS, "/price_map.png"), price_map)
```

## Network of hosts
```{r}
multi_hosts = d1 %>% filter(ID %in% topID) %>% sdf_collect()

lat = multi_hosts %>% summarise(lat = mean(latitude)) %>% as.numeric()
lon = multi_hosts %>% summarise(lon = mean(longitude)) %>% as.numeric()

nyc = get_googlemap(center = c(lon = lon, lat = lat),
                     zoom = 12, scale = 2,
                     maptype = "roadmap",
                     color = "color")

ggmap(nyc) +
  geom_line(data = multi_hosts,aes(x = longitude, y = latitude, color = factor(ID)))+
  geom_point(data = multi_hosts,aes(x = longitude, y = latitude, color = factor(ID)),size=0.3)+
  guides(color=F)+
  theme_void()+
  scale_color_manual(values = rainbow(length(unique(multi_hosts$ID))))

cols = rep("#002a75", 36079)
(network = ggplot(d1,aes(x = longitude, y = latitude, color=factor(ID)))+
  geom_line(aes(alpha=price))+
  geom_point(aes(alpha=price), size=0.01)+
  guides(color=F, alpha=F)+
  theme_void()+
  scale_color_manual(values = cols)+
  theme(plot.background = element_rect(fill="black", color="black")))

golden_ratio = (1+sqrt(5))/2
n = 7
ggsave(paste0(VISUALIZATIONS, "/network_full.png"), network, dpi=600, height = n, width = n*golden_ratio)
```

```{r}
write = sdf_collect(d1)
readr::write_csv(write, "data/cleaned_data.csv")
```












