---
title: "03 Modelling"
author: "Dominik Klepl"
date: "2/7/2020"
output: html_document
---

```{r setup}
library(rsparkling)
library(sparklyr)
library(dplyr)
library(ggplot2) #for plots
library(ggthemes)
library(magrittr) #pipes support
library(h2o)
library(patchwork)

sc = spark_connect(master = "local", version = "2.2.1")

DATA_PATH = "data/cleaned_data.csv"
VISUALIZATIONS = "documentation/figures"
data = spark_read_csv(sc = sc, name = "data", path = DATA_PATH)
```

### Set schema
```{r}
data = data %>% 
  mutate(ID = as.numeric(ID)) %>%
  mutate(latitude = as.numeric(latitude)) %>%
  mutate(longitude = as.numeric(longitude)) %>%
  mutate(price = as.numeric(price)) %>%
  mutate(minimum_nights = as.numeric(minimum_nights)) %>%
  mutate(number_of_reviews = as.numeric(number_of_reviews)) %>%
  mutate(reviews_per_month = as.numeric(reviews_per_month)) %>%
  mutate(listing_host = as.numeric(listing_host)) %>%
  mutate(availability = as.numeric(availability)) %>%
  na.omit() %>%
  select(-desc, -last_review, -ID)
```

# Data transformation

### Price
```{r}
transformed = data %>%
  filter(price < 1000) %>%
  mutate(price = log(price))

(price_dens = ggplot(transformed, aes(price))+
  geom_density())
```

### Minimum nights
```{r}
transformed = transformed %>%
  mutate(minimum_nights = log1p(minimum_nights))

(minimum_dens = ggplot(transformed)+
  geom_density(aes(minimum_nights)))
```

### N of reviews
```{r}
#feature engineer - no reviews
transformed = transformed %>%
  mutate(no_reviews = case_when(number_of_reviews==0 ~ 1,
                                number_of_reviews!=0 ~ 0))

transformed = transformed %>%
  mutate(number_of_reviews = log1p(number_of_reviews))

(n_reviews_dens = ggplot(transformed,aes(number_of_reviews))+
  geom_density())
```

### Reviews per month
```{r}
transformed = transformed %>%
  mutate(reviews_per_month = log1p(reviews_per_month))

(reviews_month_dens = ggplot(transformed, aes(reviews_per_month))+
  geom_density())
```

### Availability
```{r}
transformed = transformed %>%
  mutate(all_year = case_when(availability >= 353 ~ 1,
                              availability < 353 ~ 0)) %>%
  mutate(low_availability = case_when(availability <= 14 ~ 1,
                              availability > 14 ~ 0))

(availability_dens = ggplot(transformed, aes(availability))+
  geom_density())
```

# H2O + Spark = Sparkling Water
Start H2O
```{r}
hc = H2OContext.getOrCreate(sc)
hc$openFlow
```

Make the data accesible to H2O
```{r}
df_h2o = hc$asH2OFrame(transformed)

df_h2o$neighbourhood = as.factor(df_h2o$neighbourhood)
df_h2o$neighbourhood_group = as.factor(df_h2o$neighbourhood_group)
df_h2o$room_type = as.factor(df_h2o$room_type)
df_h2o$no_reviews = as.factor(df_h2o$no_reviews)
df_h2o$all_year = as.factor(df_h2o$all_year)
df_h2o$low_availability = as.factor(df_h2o$low_availability)
```

## Train-test split
We'll use 80% for training and 20% for testing.
```{r}
sets = h2o.splitFrame(df_h2o, ratios = 0.8)

train = sets[[1]]
test = sets[[2]]

cat("Training:", nrow(sets[[1]]), "\nTesting: ", nrow(sets[[2]]))
```

#### Variables for models
```{r}
x = colnames(df_h2o)[-6]
y = colnames(df_h2o)[6]
n_folds = 10
```


# Model training

##Generalized linear model
The function implements random search for best lambda
```{r}
glm_fit <- h2o.glm(x = x,
                   y = y,
                   training_frame = train,
                   nfolds = n_folds,
                   lambda_search = T,
                   early_stopping = T
               )

#save results
(CV_glm = glm_fit@model$cross_validation_metrics_summary %>% select(mean, sd))

(glm_test = h2o.performance(glm_fit, newdata = test))
```

##Random forest
```{r}
rf_par = list(ntrees = seq(10,80,2),
              max_depth = seq(4,16,1))
strategy = list(strategy = "RandomDiscrete", 
                         max_models = 50)


forests = h2o.grid("randomForest",
                   grid_id = "forest_grid",
                   x = x,
                  y = y,
                  training_frame = train,
                  nfolds = n_folds,
                  categorical_encoding = "AUTO",
                  hyper_params = rf_par,
                  search_criteria = strategy)

forests_summary = h2o.getGrid(grid_id = "forest_grid",
                              sort_by = "mse",
                              decreasing = T)
forests_summary

#get the best random forest
best_forest = h2o.getModel(forests_summary@model_ids[[1]])

(CV_forest = best_forest@model$cross_validation_metrics_summary %>% select(mean, sd))

#compute performance on testing
(forest_test = h2o.performance(best_forest, newdata = test))
```


##XGBoosts
```{r}
automl = h2o.automl(x = x,
                    y = y,
                    training_frame = train,
                    nfolds = 10,
                    stopping_metric = "AUTO",
                    project_name = "automl",
                    max_runtime_secs = 4000
                    )
```

# Model evaluation
Let's compare the performance of all 3 trained models. The selection metric is MSE.
```{r}
#rbind the results together
results = rbind(CV_glm[1,],
                CV_forest[1,])
results$test = c(glm_test@metrics$MSE, forest_test@metrics$MSE)
results$model = factor(c("Elastic Net", "Random Forest"))

results$mean = as.numeric(results$mean)
results$sd = as.numeric(results$sd)
results$test = as.numeric(results$test)


(results_plot = ggplot(results, aes(x = model, y = mean, fill = model))+
  geom_col()+
  geom_errorbar(aes(ymin = mean-2.5*sd, ymax = mean+2.5*sd))+
  geom_point(aes(y = test), size = 3)+
  theme_few()+
  scale_fill_tableau()+
  guides(fill = F)+
  labs(subtitle = "X denotes score on testing data",
       x = "",
       y = "MSE")+
  theme(plot.subtitle = element_text(hjust = 0.5)))

ggsave(paste0(VISUALIZATIONS, "/model_results.png"), results_plot)
```

# Model interpretation

## Feature importance plots
```{r}
importance_glm = h2o.varimp(glm_fit)
importance_forest = h2o.varimp(best_forest)

cols = rep("#446794", 10)

(imp_plot_glm = ggplot(importance_glm[1:10,], aes(x = reorder(variable, scaled_importance), y = scaled_importance, fill = factor(variable)))+
  geom_col()+
  theme_few()+
  guides(fill = F)+
  labs(x = "",
       y = "",
       title = "Feature importance for GLM")+
  coord_flip()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = cols))

ggsave(paste0(VISUALIZATIONS, "/imp_glm.png"), imp_plot_glm)

(imp_plot_rf = ggplot(importance_forest[1:10,], aes(x = reorder(variable, scaled_importance), y = scaled_importance, fill = factor(variable)))+
  geom_col()+
  theme_few()+
  guides(fill = F)+
  labs(x = "",
       y = "",
       title = "Feature importance for Random Forest")+
  coord_flip()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = cols))

ggsave(paste0(VISUALIZATIONS, "/imp_rf.png"), imp_plot_rf)
```

## Partial Dependence plots
```{r}
numeric_cols = colnames(transformed)[c(3:4,7:11)]

#compute partial plots for all numeric features
glm_partial = h2o.partialPlot(glm_fit, data = train, cols = numeric_cols, nbins = 30)
forest_partial = h2o.partialPlot(best_forest, data = train, cols = numeric_cols, nbins = 30)

#transform to normal scale
glm_partial[[3]] = mutate_all(glm_partial[[3]], exp)
glm_partial[[4]] = mutate_all(glm_partial[[4]], exp)
glm_partial[[5]] = mutate_all(glm_partial[[5]], exp)
forest_partial[[3]] = mutate_all(forest_partial[[3]], exp)
forest_partial[[4]] = mutate_all(forest_partial[[4]], exp)
forest_partial[[5]] = mutate_all(forest_partial[[5]], exp)

glm_partial[[1]][,2:4] = mutate_all(glm_partial[[1]][,2:4], exp)
glm_partial[[2]][,2:4] = mutate_all(glm_partial[[2]][,2:4], exp)
glm_partial[[6]][,2:4] = mutate_all(glm_partial[[6]][,2:4], exp)
glm_partial[[7]][,2:4] = mutate_all(glm_partial[[7]][,2:4], exp)
forest_partial[[1]][,2:4] = mutate_all(forest_partial[[1]][,2:4], exp)
forest_partial[[2]][,2:4] = mutate_all(forest_partial[[2]][,2:4], exp)
forest_partial[[6]][,2:4] = mutate_all(forest_partial[[6]][,2:4], exp)
forest_partial[[7]][,2:4] = mutate_all(forest_partial[[7]][,2:4], exp)

#combine data to one df
glm_partials = data.frame()

for (i in 1:7){
  partial = glm_partial[[i]]
  partial$variable = rep(colnames(partial)[1], 30)
  colnames(partial)[1] = "x"
  glm_partials = rbind(glm_partials, partial)
}

forest_partials = data.frame()

for (i in 1:7){
  partial = forest_partial[[i]]
  partial$variable = rep(colnames(partial)[1], 30)
  colnames(partial)[1] = "x"
  forest_partials = rbind(forest_partials, partial)
}

readr::write_csv(glm_partials, "data/glm_partial.csv")
readr::write_csv(forest_partials, "data/forest_partial.csv")
```

Plot the partial plots
```{r}
var = "latitude"

plot_partial = function(var, data, model) {
  plot = data %>% filter(variable==var) %>%
    ggplot(aes(x = x, y = mean_response, 
             ymin = mean_response-std_error_mean_response,
             ymax = mean_response+std_error_mean_response))+
    geom_line(color = "#4E79A7")+
    geom_ribbon(alpha = 0.2, fill = "#F28E2B")+
    theme_few()+
    labs(x = "",
       y = "price",
       title = toupper(var))+
    theme(plot.title = element_text(hjust = 0.5))
  
  ggsave(paste0(VISUALIZATIONS, "/", var, "_" ,model, ".png"), plot)
  return(plot)
}

(latitude_glm = plot_partial("latitude", glm_partials, "glm"))
(longitude_glm = plot_partial("longitude", glm_partials, "glm"))
(minimum_glm = plot_partial("minimum_nights", glm_partials, "glm"))
(reviewsn_glm = plot_partial("number_of_reviews", glm_partials, "glm"))
(reviewsm_glm = plot_partial("reviews_per_month", glm_partials, "glm"))
(listhost_glm = plot_partial("listing_host", glm_partials, "glm"))
(avail_glm = plot_partial("availability", glm_partials, "glm"))

(latitude_forest = plot_partial("latitude", forest_partials, "forest"))
(longitude_forest = plot_partial("longitude", forest_partials, "forest"))
(minimum_forest = plot_partial("minimum_nights", forest_partials, "forest"))
(reviewsn_forest = plot_partial("number_of_reviews", forest_partials, "forest"))
(reviewsm_forest = plot_partial("reviews_per_month", forest_partials, "forest"))
(listhost_forest = plot_partial("listing_host", forest_partials, "forest"))
(avail_forest = plot_partial("availability", forest_partials, "forest"))
```


